---
title: "Lab 5 - Data Wrangling"
author: "Eugene Nguyen"
date: "9/24/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load Libraries
```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(data.table)
library(leaflet)
```

### Load Station Data
```{r}
# Download the data
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```

### Load Met Data
```{r}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
```

### Merge Station & Met
```{r}
df <- merge(
  # Data
  x     = met,      
  y     = stations, 
  # List of variables to match
  by.x  = "USAFID",
  by.y  = "USAF", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  )
```

# Question 1: Representative station for the US
What is the median station in terms of temperature, wind speed, and atmospheric pressure? Look for the three weather stations that best represent continental US using the quantile() function. Do these three coincide?

Knit the document, commit your changes, and Save it on GitHub. Don’t forget to add README.md to the tree, the first time you render it.
```{r}
# Average Station for temperature, wind speed, and atmospheric pressure
# Tidyverse Method
# station_avg <- df %>%
#   group_by(USAFID) %>%
#   summarise(temp_avg = mean(temp, na.rm = TRUE)
#             , wind.speed_avg = mean(wind.sp, na.rm = TRUE)
#             , atm.press_avg = mean(atm.press, na.rm = TRUE)
#             )

# data.table method
# Average Station for temperature, wind speed, and atmospheric pressure
stations_avg <- df[,.(
  temp = mean(temp, na.rm = TRUE),
  wind.sp = mean(wind.sp, na.rm = TRUE),
  atm.press = mean(atm.press, na.rm = TRUE)
), by = USAFID]
```


```{r}
# Identify quantiles per variable
medians <- stations_avg[,.(
  temp_50 = quantile(temp, probs = .5, na.rm = TRUE),
  wind.sp_50 = quantile(wind.sp, probs = .5, na.rm = TRUE),
  atm_press_50 = quantile(atm.press, probs = .5, na.rm = TRUE)
)]

medians
```

```{r}
# Find stations closest to these values
# Using which.min() function 
# Median temp
stations_avg[, temp_dist := abs(temp - medians$temp_50)]
median_temp_station <- stations_avg[order(temp_dist)[1]]
median_temp_station

# Median wind speed
stations_avg[, wind.sp_dist := abs(wind.sp - medians$wind.sp_50)]
median_windsp_station <- stations_avg[order(wind.sp_dist)[1]]
median_windsp_station

# Median atm press
stations_avg[, atm.press_dist := abs(atm.press - medians$atm_press_50)]
median_atmpress_station <- stations_avg[order(atm.press_dist)[1]]
median_atmpress_station
```
The station that is closest to the median temperature is `r median_temp_station$USAFID`.
The station that is closest to the median wind speed is `r median_windsp_station$USAFID`.
The station that is closest to the median wind speed is `r median_atmpress_station$USAFID`.


# Question 2: Representative station per state
Just like the previous question, you are asked to identify what is the most representative, the median, station per state. This time, instead of looking at one variable at a time, look at the euclidean distance. If multiple stations show in the median, select the one located at the lowest latitude.

Knit the doc and save it on GitHub.
```{r}
# Merge state information
stations_avg2 <- merge(x = stations_avg, y = stations, by.x = 'USAFID', by.y = "USAF", all.x = TRUE, all.y = FALSE)

```

```{r}
# Get the medians by state for temperature, wind speed, and atm press
stations_avg2[, temp_50 := quantile(temp, probs = .5, na.rm = TRUE), by = STATE]
stations_avg2[, windsp_50 := quantile(wind.sp, probs = .5, na.rm = TRUE), by = STATE]
stations_avg2[, atmpress_50 := quantile(atm.press, probs = .5, na.rm = TRUE), by = STATE]
```

```{r}
# Calculate Euclidean Distance
stations_avg2[, eudist := sqrt(
  (temp - temp_50)^2 + (wind.sp - windsp_50)^2
                              )]
```

```{r}
# Lowest euclidean distance by State
stations_avg2[ , .SD[which.min(eudist)], by = STATE]
```
The station with the lowest euclidean distance between temperature and wind speed is `r median_atmpress_station$USAFID`.


# Question 3: In the middle?
For each state, identify what is the station that is closest to the mid-point of the state. Combining these with the stations you identified in the previous question, use leaflet() to visualize all ~100 points in the same figure, applying different colors for those identified in this question.

Knit the doc and save it on GitHub.
```{r}
# Calculate midpoint by State
stations_avg2[, midpoint := sqrt(
  ((temp - temp_50)^2 + (wind.sp - windsp_50)^2) / 2
                              )]
```

```{r}
# Lowest midpoint by State
map <- stations_avg2[ , .SD[which.min(midpoint)], by = STATE]
```

```{r}
# Create table to map lat/lon
hashtable <- df %>%
  select(USAFID, lat, lon)

hashtable <- distinct(hashtable, USAFID, .keep_all = TRUE)
```

```{r}
# Merge lat/lon data
# Merge state information
map2 <- merge(x = map, y = hashtable, by.x = 'USAFID', by.y = "USAFID", all.x = TRUE, all.y = FALSE)
```



```{r}
# Create leaflet map
mp.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=map2$midpoint)

# Leadlet map
tempmap <- leaflet(map2) %>% 
  # The looks of the Map
  addProviderTiles('CartoDB.Positron') %>% 
  # Some circles
  addCircles(
    lat = ~lat, lng=~lon,
                                                  # HERE IS OUR PAL!
    label = ~paste0(round(temp,2), ' C'), color = ~ mp.pal(midpoint),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=mp.pal, values=map2$midpoint,
          title='Midpoint by State', opacity=1)

tempmap
```


# Question 4: Means of means
Using the quantile() function, generate a summary table that shows the number of states included, average temperature, wind-speed, and atmospheric pressure by the variable “average temperature level,” which you’ll need to create.

Start by computing the states’ average temperature. Use that measurement to classify them according to the following criteria:

low: temp < 20
Mid: temp >= 20 and temp < 25
High: temp >= 25

```{r}
# Create
df[, state_temp := mean(temp, na.rm = TRUE), by = STATE]
df[, temp_cat := fifelse(
  state_temp < 20, "low-temp",
  fifelse(state_temp < 25, "mid-temp", "high-temp")
  )]
```



Once you are done with that, you can compute the following:

Number of entries (records),
Number of NA entries,
Number of stations,
Number of states included, and
Mean temperature, wind-speed, and atmospheric pressure.
All by the levels described before.

```{r}
# Summary table
df[, .(
  N_entries = .N,
  N_stations = length(unique(USAFID)),
  N_missing = sum(is.na(.SD)),
  N_states = length(unique(STATE)),
  mean_temperature = mean(temp, na.rm = TRUE),
  mean_windspeed = mean(wind.sp, na.rm = TRUE),
  mean_atmpress = mean(atm.press, na.rm = TRUE)
), by = temp_cat]
```



