---
title: "Lab 3 - Exploratory Data Analysis"
author: "Gene Nguyen"
date: "9/10/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Libraries
```{r message = FALSE}
# Load packages
library(tidyverse)
library(leaflet)
library(lubridate)
```

Set Workspace
```{r}
setwd("C:/Users/Eugen/Desktop/PM566/PM565-labs/lab3")
```

### 1. Read in Data
```{r}
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
met <- data.table::fread("met_all.gz")
```

### 2. Check the dimensions, headers, footers. How many columns, rows are there?
```{r}
# Dimension
dim(met)
# Rows
nrow(met)
# Columns
ncol(met)
# Header
head(met)
# Footer
tail(met)
```

There are 30 columns and 2377343 in the met dataset.

### 3. Take a look at the variables
```{r}
str(met)
```


### 4. Take a closer look at key variables
```{r}
table(met$year)
table(met$day)
table(met$hour)
summary(met$temp)
summary(met$elev)
```

Elevation looks like <- it has missing values (max = 9999). Replace with NA
```{r}
met[met$elev == 9999.0] <- NA
summary(met$elev)
```

At what elevation is the highest weather station?
- The weather station with highest elevation is 4113 meters.
- This is after replacing 9999.0 values with the appropriate code for “missing”, which is “NA”.

There is a minimum temperature of -40...doesn't seem plausible so we should remove
```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```

### 5. Check the data against an external data source.
```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```

# 6. Calculate summary statistics
```{r}
elev <- met[elev==max(elev)]
summary(elev)

# Correlation (temp vs. wind)
cor(elev$temp, elev$wind.sp, use="complete")
# Correlation (temp vs time)
cor(elev$temp, elev$hour, use="complete")
# Correlation (wind vs time)
cor(elev$wind.sp, elev$day, use="complete")
# Correlation (wind vs time)
cor(elev$wind.sp, elev$hour, use="complete")
# Correlation (wind vs time)
cor(elev$temp, elev$day, use="complete")
```

# 7. Exploratory Graphs
```{r}
# Histogram of Elevation
hist(met$elev, breaks=100)

# Histogram of Temperature
hist(met$temp)

# Histogram of Wind
hist(met$wind.sp)
```

Wind speed and elevation look skewed. Possibly look to log transform in the future.

Location - Weather station with the highest elevation
```{r}
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>%
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

Look at the time series of temperature and wind speed at this location.
For this we will need to create a date-time variable for the x-axis.
```{r}
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)

# Order
elev <- elev[order(date)]
head(elev)
```

Time Series: Temperature
```{r}
plot(elev$date, elev$temp, type='l')
```

There appears to be seasonality with temperature. It seems to steadily fluctuate every other day.

Time Series: Wind
```{r}
plot(elev$date, elev$wind.sp, type='l')
```

There appears to be seasonality with wind, but much more erratic. There is more chaotic fluctuation when compared to temperature.
