---
title: "Lab 6 - Text Mining"
author: "Eugene Nguyen"
date: "10/5/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, warning = FALSE}
# Load Libraries
library(tidyverse)
library(tidytext)
library(tibble)
```


## Download The Data
```{r}
fn <- "mtsamples.csv"
if (!file.exists(fn))
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv", destfile = fn)

mtsamples <- read.csv(fn)
mtsamples <- as.tibble(mtsamples)
```
Take a look at the dataset
```{r}
head(mtsamples)
```


## Question 1: What specialties do we have?
We can use count() from dplyr to figure out how many different catagories do we have? Are these catagories related? overlapping? evenly distributed?
```{r}
specialties <- 
  mtsamples %>%
  count(medical_specialty) %>%
  arrange(desc(n)) %>%
  top_n(n, 15) %>%
  knitr::kable()

specialties
```
There are `r nrow(specialties)` specialties.

Let's examine the distribution:
```{r}
# Histogram for Medical Speciality
ggplot(mtsamples, aes(x = medical_specialty)) +
  geom_histogram(stat = "count") +
  coord_flip()
```
These categories are not evenly distributed.

## Question 2
Tokenize the the words in the transcription column
Count the number of times each token appears
Visualize the top 20 most frequent words
Explain what we see from this result. Does it makes sense? What insights (if any) do we get?
```{r}
# Tokenize the Transcription Column
transcription_20 <- 
  mtsamples %>%
  unnest_tokens(token, transcription) %>%
  count(token, sort = TRUE) %>%
  top_n(20, n)

transcription_20 %>%
  knitr::kable()
```
```{r}
# Visualize the top 20
ggplot(transcription_20, aes(x = token, y = n)) +
  geom_bar(stat = "identity") +
  coord_flip()
```
There are many unnecessary words in the histogram. These are also known as "stop words". There are available functions to remove them.

## Question 3
Redo visualization but remove stopwords before
Bonus points if you remove numbers as well
What do we see know that we have removed stop words? Does it give us a better idea of what the text is about?
```{r}
# Remove Stop Words
transcription_20_sw <- 
  mtsamples %>%
  unnest_tokens(output = word, input = transcription) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = "word") %>%
  top_n(20)

transcription_20_sw %>%
  knitr::kable()
```
```{r}
# Visualize the top 20 without the stop words
ggplot(transcription_20_sw, aes(x = word, y = n)) +
  geom_bar(stat = "identity") +
  coord_flip()
```
The graph looks a lot better, but there are numbers that we should remove.

```{r}
# Remove stop Words AND numbers
transcription_20_sw2 <- 
  mtsamples %>%
  unnest_tokens(output = word, input = transcription) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = "word") %>%
  filter(!grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(20)

transcription_20_sw2 %>%
  knitr::kable()
```

```{r}
# Visualize the top 20 without the stop words AND numbers
ggplot(transcription_20_sw2, aes(x = word, y = n)) +
  geom_bar(stat = "identity") +
  coord_flip()
```


## Question 4
repeat question 2, but this time tokenize into bi-grams. how does the result change if you look at tri-grams?
```{r}
# Tokenize the Transcription Column into bi-grams/tri-grams
# bi-gram
transcription_bg <- 
  mtsamples %>%
  unnest_ngrams(ngram, transcription, n = 2) %>%
  count(ngram, sort = TRUE) %>%
  top_n(20, n)

# tri-gram
transcription_tg <- 
  mtsamples %>%
  unnest_ngrams(ngram, transcription, n = 3) %>%
  count(ngram, sort = TRUE) %>%
  top_n(20, n)

transcription_bg %>%
  knitr::kable()

transcription_tg %>%
  knitr::kable()
```

```{r}
# Visualize the top 20 bi-grams
ggplot(transcription_bg, aes(x = ngram, y = n)) +
  geom_bar(stat = "identity") +
  coord_flip()

# Visualize the top 20 tri-grams
ggplot(transcription_tg, aes(x = ngram, y = n)) +
  geom_bar(stat = "identity") +
  coord_flip()
```
It looks like the tri-gram is more evenly distributed, with the exception of "the patient was" appearing more times than the others.

## Question 5
Using the results you got from questions 4. Pick a word and count the words that appears after and before it.

Let's take a look at the word "patient" since it shows up frequently.
```{r}
# Create a tri-gram and separate into 3 columns
# Also remove stop words & numbers
# Before
mtsamples %>%
  unnest_ngrams(ngram, transcription, n = 3) %>%
  separate(ngram, into = c("before", "word", "after"), sep = " ") %>%
  select(before, word, after) %>%
  filter(word == "patient") %>%
  count(before, sort = TRUE) %>%
  anti_join(stop_words, by = c("before" = "word")) %>%
  filter(!grepl(pattern = "^[0-9]+$", x = before)) %>%
  top_n(20) %>%
  knitr::kable()

# After
# Before
mtsamples %>%
  unnest_ngrams(ngram, transcription, n = 3) %>%
  separate(ngram, into = c("before", "word", "after"), sep = " ") %>%
  select(before, word, after) %>%
  filter(word == "patient") %>%
  count(after, sort = TRUE) %>%
  anti_join(stop_words, by = c("after" = "word")) %>%
  filter(!grepl(pattern = "^[0-9]+$", x = after)) %>%
  top_n(20) %>%
  knitr::kable()
```


## Question 6
Which words are most used in each of the specialties. you can use group_by() and top_n() from dplyr to have the calculations be done within each specialty. Remember to remove stopwords. How about the most 5 used words?
```{r}
mtsamples %>%
  unnest_tokens(word, input = transcription) %>%
  group_by(medical_specialty) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = "word") %>%
  filter(!grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(5) %>%
  arrange(medical_specialty, desc(n)) %>%
  knitr::kable()
```


## Question 7 - extra
Find your own insight in the data:

Ideas:

Interesting ngrams
See if certain words are used more in some specialties then others

Let's see the top 5 associated sample names with each medical specialty.
```{r}
mtsamples %>%
  unnest_tokens(word, input = sample_name) %>%
  group_by(medical_specialty) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = "word") %>%
  filter(!grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(5) %>%
  arrange(medical_specialty, desc(n)) %>%
  knitr::kable()
```
Very Interesting! Hospice caught my attention with "hospice", "visit", and "cancer" being the top 3 within this group.
